### 2023-06-02 manuscript/manuscript.qmd
Added first two paragraphs to Introduction

---
title: Climate predictor selection and variable reduction methods influence MaxEnt model performance and predictions of climatic suitability 
author:
  - name: Clarke van Steenderen
    email: vsteenderen@gmail.com
    affiliations: 
        - id: some-tech
          name: Center for Biological Control, Department of Zoology and Entomology
          department: Rhodes University
          address: 
          city: Makhanda
          state: South Africa
          postal-code: 6140
    attributes:
        corresponding: false
  - name: Guy F. Sutton
    email: g.sutton@ru.ac.za
    affiliations:
        - id: another-u
          name: Center for Biological Control, Department of Zoology and Entomology
          department: Rhodes University
          address: 
          city: Makhanda
          state: South Africa
          postal-code: 6140
    attributes:
        corresponding: true
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - keyword1
  - keyword2
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Ecological Modelling
      formatting: preprint
      model: 3p
      layout: twocolumn
      cite-style: authoryear
      highlights:
        - Highlight 1
        - Highlight 2 
        - Highlight 3
---

# Introduction

Ecological models are important tools to aid the development and implementation of environmental policies and management programmes [@Addison2013; @Schuwirth2019; @Sutton2022]. These models are used for conservation planning [@Guisan2013], predicting the establishment and spread of invasive species [@Martin2020], implementing biological control programmes [@Sutton2019b; @Mukherjee2021], and forecasting species responses to environmental change [@Bocedi2014], amongst other applications. Species distribution models (SDM’s) are an example of ecological models that have become increasingly popular in recent years [@Elith2009]. SDM’s typically take the form of correlative or mechanistic models that correlate species presence/absences (or pseudo-absences) to environmental covariates to identify suitable climatic conditions for the study taxon [@Elith2011]. The Maximum Entropy species distribution model (hereafter ‘MaxEnt’) is amongst the most popular methods for climate modelling studies and has been shown to perform well compared to alternative modelling algorithms [@Wisz2008; @Phillips2017]. It uses maximum entropy to distinguish between environmental conditions where the focal taxon is present from environmental conditions at sites without confirmed presence records for the taxon [@Elith2011].

In recent years, a number of studies have investigated and demonstrated that computational choices made during the model building process can have a significant influence on resulting model outputs and inferences drawn [@Warren2011; @Webber2011; @Shcheglovitova2013; @Boria2017; @Sutton2022]. Despite its importance in the model building process, covariate selection methods have received considerably little attention to date [but see @Austin2011; @Fourcade2018; @Adde2023], whereby covariate selection refers to "identify[ing] the best subset of covariates out of a panel of many candidates, both from an ecological and statistical perspective [see @Adde2023 and references therein].

### 2023-06-02 manuscript/manuscript.tex
Added first two paragraphs to Introduction

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  authoryear,
  preprint,
  3p,
  twocolumn]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\usepackage{float}
\makeatletter
\let\oldlt\longtable
\let\endoldlt\endlongtable
\def\longtable{\@ifnextchar[\longtable@i \longtable@ii}
\def\longtable@i[#1]{\begin{figure}[H]
\onecolumn
\begin{minipage}{0.5\textwidth}
\oldlt[#1]
}
\def\longtable@ii{\begin{figure}[H]
\onecolumn
\begin{minipage}{0.5\textwidth}
\oldlt
}
\def\endlongtable{\endoldlt
\end{minipage}
\twocolumn
\end{figure}}
\makeatother
\journal{Ecological Modelling}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Climate predictor selection and variable reduction methods influence MaxEnt model performance and predictions of climatic suitability},
  pdfauthor={Clarke van Steenderen; Guy F. Sutton},
  pdfkeywords={keyword1, keyword2},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Climate predictor selection and variable reduction methods
influence MaxEnt model performance and predictions of climatic
suitability}
\author[1]{Clarke van Steenderen%
%
}
 \ead{vsteenderen@gmail.com} 
\author[1]{Guy F. Sutton%
\corref{cor1}%
}
 \ead{g.sutton@ru.ac.za}

\affiliation[1]{organization={Center for Biological Control, Department
of Zoology and Entomology, Rhodes
University},city={Makhanda},postcode={6140},postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{abstract}
This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing
elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget
velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris
sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta
egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet
auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo
molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae.
Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at
tempus elit.
\end{abstract}

\begin{highlights}
\item Highlight 1\item Highlight 2\item Highlight 3
\end{highlights}

\begin{keyword}
    keyword1 \sep 
    keyword2
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Ecological models are important tools to aid the development and
implementation of environmental policies and management programmes
\citep{Addison2013, Schuwirth2019, Sutton2022}. These models are used
for conservation planning \citep{Guisan2013}, predicting the
establishment and spread of invasive species \citep{Martin2020},
implementing biological control programmes
\citep{Sutton2019b, Mukherjee2021}, and forecasting species responses to
environmental change \citep{Bocedi2014}, amongst other applications.
Species distribution models (SDM's) are an example of ecological models
that have become increasingly popular in recent years \citep{Elith2009}.
SDM's typically take the form of correlative or mechanistic models that
correlate species presence/absences (or pseudo-absences) to
environmental covariates to identify suitable climatic conditions for
the study taxon \citep{Elith2011}. The Maximum Entropy species
distribution model (hereafter `MaxEnt') is amongst the most popular
methods for climate modelling studies and has been shown to perform well
compared to alternative modelling algorithms
\citep{Wisz2008, Phillips2017}. It uses maximum entropy to distinguish
between environmental conditions where the focal taxon is present from
environmental conditions at sites without confirmed presence records for
the taxon \citep{Elith2011}.

In recent years, a number of studies have investigated and demonstrated
that computational choices made during the model building process can
have a significant influence on resulting model outputs and inferences
drawn
\citep{Warren2011, Webber2011, Shcheglovitova2013, Boria2017, Sutton2022}.
Despite its importance in the model building process, covariate
selection methods have received considerably little attention to date
\citep[but see][]{Austin2011, Fourcade2018, Adde2023}, whereby covariate
selection refers to ``identify{[}ing{]} the best subset of covariates
out of a panel of many candidates, both from an ecological and
statistical perspective \citep[see][ and references therein]{Adde2023}.

  \bibliography{bibliography.bib}

\end{document}

### 2023-06-02 manuscript/manuscript.qmd
Add Acknowledgements and data availability statement

@@
 
-# Introduction
+# 1. Introduction
 
@@
 Despite its importance in the model building process, covariate selection methods have received considerably little attention to date [but see @Austin2011; @Fourcade2018; @Adde2023], whereby covariate selection refers to "identify[ing] the best subset of covariates out of a panel of many candidates, both from an ecological and statistical perspective [see @Adde2023 and references therein].
+
+**Next paragraphs**
+
+- Review current methods used to select variables, e.g. R2, VIF, PCA, expert opinion, and most recently, automated selection / model-based selection.
+
+- Aims of this paper are to show how variable selection can influence model outputs and inferences.
+
+# 2. Methods and materials
+
+# 3. Results
+
+# 4. Discussion
+
+# Declaration of Competing Interest
+
+The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.
+
+# Data availability
+
+All data and code required to reproduce the analyses are available in a public GitHub repository: https://github.com/guysutton/MS_climate_variable_selection_sdm
+
+# Acknlowedgements
+
+CVS and GFS acknowledge funding from the South African Working for Water (WfW) programme of the Department of Forestry, Fisheries and the Environment: Natural Resource Management Programmes (DFFE: NRMP).
+Funding was also provided by the South African Research Chairs Initiative of the Department of Science and Technology and the National Research Foundation (NRF) of South Africa.
+Any opinions, finding, conclusions or recommendations expressed in this material are those of the authors and the NRF does not accept any liability in this regard.
+
+# Supplementary materials
+
+# References

### 2023-06-02 manuscript/manuscript.qmd
Add methods text bulk

@@
 
+### 2.1. Species occurrence records
+
+### 2.2. Environmental predictors
+
+Climate data were obtained by downloading the standard set of 19 bioclimatic variables from the WorldClim ver.
+2.1 database (Fick and Hijmans, 2017) (data available at: www.worldclim.org/download.
+html).
+This dataset is representative of annual and seasonal means and variation of temperature and precipitation metrics averaged over the 1950–2000 time period (current climate) at a 2.5 min resolution.
+These variables have been shown to effectively model the climatic suitability for non-native insects (e.g., Trethowan et al., 2011).
+
+### 2.3. Model calibration
+
+MaxEnt (ver. 3.4.3) was implemented in the ‘ dismo’ R package (Hijmans et al., 2021 ).
+
+Given that MaxEnt is a presence/pseudo-absence modelling algorithm, model calibration requires a user-defined geographic background to sample the climate of representative grid cells where the focal species is assumed to be absent (i.e., background points or pseudo-absences).
+Background definition can have a significant effect on model output (VanDerWal et al., 2009).
+The background should ideally represent the geographic areas available to the focal species, omitting areas where species absence is due to historical factors, dispersal constraints and/or biotic interactions (Sanin and Anderson, 2018).
+Following Webber et al.
+(2011), we defined the model background using the Koppen-Geiger climate classification system (available at: http://koeppen-geiger.vu-w ien.ac.at).
+Only Koppen-Geiger climate zones that contained at least one native-range occurrence record for D.
+rubiformis in Australia were used as the background area from which background points were drawn for model calibration (Fig.
+1a).
+The Koppen-Geiger climate zones were intersected with the occurrence records using the ‘ raster’ R package (Hijmans, 2022).
+We randomly sampled 10 000 points (the default number used for Maxent; Merow et al.
+(2013) ) from within this background definition using the ‘dismo’ R package (Hijmans et al., 2021).
+
+MaxEnt models were parameterised with default settings for multiple parameters, including: convergence = 105, maximum number of iterations = 500 and prevalence = 0.5.
+The ‘fade by clamping’ option was selected to prevent extrapolation well outside the range of climatic values in the model training area (Philips et al., 2017 ).
+Model predictions were obtained using the ‘logistic output’ to create continuous climatic suitability raster layers scaled between 0 (climatically unsuitable) and 1 (climatically suitable).
+
+### 2.4. Model evaluation
+
+Model tuning experiments were applied to the native-range MaxEnt models to derive within-sample evaluation metrics to guide the selection of optimal MaxEnt parameter configurations (feature classes and regularisation multipliers).
+Optimised parameter configurations would then be used to refit the MaxEnt models before being projected into a novel geographic region and making projections of climatic suitability for D.
+rubiformis .
+Model tuning was performed by building MaxEnt models with varying (1) feature class combinations (H = Hinge only, L = Linear only, LQ = Linear and Quadratic and LQH = Linear, Quadratic and Hinge features) and (2) regularisation multipliers (1:8).
+In total, 32 MaxEnt models were specified.
+Native-range model performance and optimal parameter configurations were assessed using 4-fold spatial block cross validation using ‘ENMeval’ (Kass et al., 2021).
+
+Optimal parameter configurations were assessed using multiple metrics that reflect different aspects of model performance.
+Four metrics were calculated, including: (1) discriminatory ability (AUCtest), (2) overfitting (AUCdiff), (3) omission rates (OR10), and (4) overall parsimony (AICc).
+The use of AUC analyses for assessing the fit of MaxEnt models has been criticised for a variety of reasons (see Lobo et al., 2008; Peterson et al., 2008 ).
+However, AUC metrics are arguably the most widely used metrics to evaluate MaxEnt model performance, and as such, we believe it is important to include them in our evaluation, and contrast the results obtained using AUC versus other metrics.
+
+We specified five final MaxEnt models, four models calibrated with FC and RM values that optimised model performance based on the metricsdiscussed below, and a MaxEnt model calibrated with default FC and RM values.
+Our intention was to compare MaxEnt model predictions and perfor mance depending on which metric was used to select optimal parameter configurations relative to the default MaxEnt settings.
+
+(1) AUC test assesses the model ’s ability to discriminate between predicted presence at withheld portions of the data used to test the model versus pseudo-absence points.
+An AUC of less than 0.8 is considered a poor model, between 0.8 and 0.9 is a fair model, between 0.9 and 0.995 a good model, and > 0.995 an excellent model (Fielding and Bell, 1997).
+Thus, higher AUCtest values indicate increased ability to discriminate between testing and background points.
+
+(2) AUC diff is the difference between AUC values calculated on training points only (AUCtrain) and AUCtest [see (1) AUCtest above for details] (Warren and Seifert, 2011).
+Thus, higher AUCdiff values indicate whether the MaxEnt model is overfit on the training data, and thus, may perform poorly when evaluated against testing points.
+
+(3) OR 10 is the 10% training omission rate (Boria et al., 2014 ).
+Overfit models have omission rates higher than the theoretical expectation for the threshold applied (Shcheglovitova and Anderson, 2013).
+As such, the OR 10 criterion selected models calibrated with MaxEnt settings which best approximated the expected 0.10 omission rate.
+Models with omission rates increasingly higher than the expected value were considered to have a higher degree of overfit (Boria et al., 2017).
+
+(4) The Akaike Information Criterion corrected for small sample sizes (AICc) criterion simultaneously scores models according to their complexity and goodness-of-fit.
+AICc was used as the primary evaluation metric as it is calculated using MaxEnt models built using the entire species occurrence dataset (i.e.
+all the occurrence points in the native-range), unlike AUC and OR10 (and numerous other metrics frequently used for model evaluation) which may be spatially biased due to the partitioning of the species occurrence dataset into training and evaluation sets (Sanin and Anderson, 2018).
+Optimal parameter configurations were determined by selecting model configurations which produced the lowest value for AICc (i.e., AICc=0; following Kass et al.
+(2021)).
+
+### 2.5. Model visualisation
+
+- Need to discuss how we map the different rasters, and maybe how we quantified the difference in suitability projections between climate layers?
+
+All modelling and statistical analyses were conducted in R ver.
+4.0.3 (R Core Team, 2020).
+All values presented in text are presentedas mean ± standard error, unless otherwise stated.
+A standardised ODMAP methods protocol (Overview, Data, Model, Assessment, and Prediction) has been completed for this study and can be found in Supplementary File S1.
+ODMAP standardises the reporting of SDM modelling studies to improve transparency and reproducibility (Zurell et al., 2020).
+
 # 3. Results

### 2023-06-02 manuscript/manuscript.tex
Render up-to-date MS

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, interior hidden, sharp corners, frame hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
 
 \hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{1. Introduction}\label{introduction}}
 
@@
 
+\textbf{Next paragraphs}
+
+\begin{itemize}
+\item
+  Review current methods used to select variables, e.g.~R2, VIF, PCA,
+  expert opinion, and most recently, automated selection / model-based
+  selection.
+\item
+  Aims of this paper are to show how variable selection can influence
+  model outputs and inferences.
+\end{itemize}
+
+\hypertarget{methods-and-materials}{%
+\section{2. Methods and materials}\label{methods-and-materials}}
+
+\hypertarget{species-occurrence-records}{%
+\subsubsection{2.1. Species occurrence
+records}\label{species-occurrence-records}}
+
+\hypertarget{environmental-predictors}{%
+\subsubsection{2.2. Environmental
+predictors}\label{environmental-predictors}}
+
+Climate data were obtained by downloading the standard set of 19
+bioclimatic variables from the WorldClim ver. 2.1 database (Fick and
+Hijmans, 2017) (data available at: www.worldclim.org/download. html).
+This dataset is representative of annual and seasonal means and
+variation of temperature and precipitation metrics averaged over the
+1950--2000 time period (current climate) at a 2.5 min resolution. These
+variables have been shown to effectively model the climatic suitability
+for non-native insects (e.g., Trethowan et al., 2011).
+
+\hypertarget{model-calibration}{%
+\subsubsection{2.3. Model calibration}\label{model-calibration}}
+
+MaxEnt (ver. 3.4.3) was implemented in the ` dismo' R package (Hijmans
+et al., 2021 ).
+
+Given that MaxEnt is a presence/pseudo-absence modelling algorithm,
+model calibration requires a user-defined geographic background to
+sample the climate of representative grid cells where the focal species
+is assumed to be absent (i.e., background points or pseudo-absences).
+Background definition can have a significant effect on model output
+(VanDerWal et al., 2009). The background should ideally represent the
+geographic areas available to the focal species, omitting areas where
+species absence is due to historical factors, dispersal constraints
+and/or biotic interactions (Sanin and Anderson, 2018). Following Webber
+et al.~(2011), we defined the model background using the Koppen-Geiger
+climate classification system (available at: http://koeppen-geiger.vu-w
+ien.ac.at). Only Koppen-Geiger climate zones that contained at least one
+native-range occurrence record for D. rubiformis in Australia were used
+as the background area from which background points were drawn for model
+calibration (Fig. 1a). The Koppen-Geiger climate zones were intersected
+with the occurrence records using the ` raster' R package (Hijmans,
+2022). We randomly sampled 10 000 points (the default number used for
+Maxent; Merow et al.~(2013) ) from within this background definition
+using the `dismo' R package (Hijmans et al., 2021).
+
+MaxEnt models were parameterised with default settings for multiple
+parameters, including: convergence = 105, maximum number of iterations =
+500 and prevalence = 0.5. The `fade by clamping' option was selected to
+prevent extrapolation well outside the range of climatic values in the
+model training area (Philips et al., 2017 ). Model predictions were
+obtained using the `logistic output' to create continuous climatic
+suitability raster layers scaled between 0 (climatically unsuitable) and
+1 (climatically suitable).
+
+\hypertarget{model-evaluation}{%
+\subsubsection{2.4. Model evaluation}\label{model-evaluation}}
+
+Model tuning experiments were applied to the native-range MaxEnt models
+to derive within-sample evaluation metrics to guide the selection of
+optimal MaxEnt parameter configurations (feature classes and
+regularisation multipliers). Optimised parameter configurations would
+then be used to refit the MaxEnt models before being projected into a
+novel geographic region and making projections of climatic suitability
+for D. rubiformis . Model tuning was performed by building MaxEnt models
+with varying (1) feature class combinations (H = Hinge only, L = Linear
+only, LQ = Linear and Quadratic and LQH = Linear, Quadratic and Hinge
+features) and (2) regularisation multipliers (1:8). In total, 32 MaxEnt
+models were specified. Native-range model performance and optimal
+parameter configurations were assessed using 4-fold spatial block cross
+validation using `ENMeval' (Kass et al., 2021).
+
+Optimal parameter configurations were assessed using multiple metrics
+that reflect different aspects of model performance. Four metrics were
+calculated, including: (1) discriminatory ability (AUCtest), (2)
+overfitting (AUCdiff), (3) omission rates (OR10), and (4) overall
+parsimony (AICc). The use of AUC analyses for assessing the fit of
+MaxEnt models has been criticised for a variety of reasons (see Lobo et
+al., 2008; Peterson et al., 2008 ). However, AUC metrics are arguably
+the most widely used metrics to evaluate MaxEnt model performance, and
+as such, we believe it is important to include them in our evaluation,
+and contrast the results obtained using AUC versus other metrics.
+
+We specified five final MaxEnt models, four models calibrated with FC
+and RM values that optimised model performance based on the
+metricsdiscussed below, and a MaxEnt model calibrated with default FC
+and RM values. Our intention was to compare MaxEnt model predictions and
+perfor mance depending on which metric was used to select optimal
+parameter configurations relative to the default MaxEnt settings.
+
+\begin{enumerate}
+\def\labelenumi{(\arabic{enumi})}
+\item
+  AUC test assesses the model 's ability to discriminate between
+  predicted presence at withheld portions of the data used to test the
+  model versus pseudo-absence points. An AUC of less than 0.8 is
+  considered a poor model, between 0.8 and 0.9 is a fair model, between
+  0.9 and 0.995 a good model, and \textgreater{} 0.995 an excellent
+  model (Fielding and Bell, 1997). Thus, higher AUCtest values indicate
+  increased ability to discriminate between testing and background
+  points.
+\item
+  AUC diff is the difference between AUC values calculated on training
+  points only (AUCtrain) and AUCtest {[}see (1) AUCtest above for
+  details{]} (Warren and Seifert, 2011). Thus, higher AUCdiff values
+  indicate whether the MaxEnt model is overfit on the training data, and
+  thus, may perform poorly when evaluated against testing points.
+\item
+  OR 10 is the 10\% training omission rate (Boria et al., 2014 ).
+  Overfit models have omission rates higher than the theoretical
+  expectation for the threshold applied (Shcheglovitova and Anderson,
+  2013). As such, the OR 10 criterion selected models calibrated with
+  MaxEnt settings which best approximated the expected 0.10 omission
+  rate. Models with omission rates increasingly higher than the expected
+  value were considered to have a higher degree of overfit (Boria et
+  al., 2017).
+\item
+  The Akaike Information Criterion corrected for small sample sizes
+  (AICc) criterion simultaneously scores models according to their
+  complexity and goodness-of-fit. AICc was used as the primary
+  evaluation metric as it is calculated using MaxEnt models built using
+  the entire species occurrence dataset (i.e.~all the occurrence points
+  in the native-range), unlike AUC and OR10 (and numerous other metrics
+  frequently used for model evaluation) which may be spatially biased
+  due to the partitioning of the species occurrence dataset into
+  training and evaluation sets (Sanin and Anderson, 2018). Optimal
+  parameter configurations were determined by selecting model
+  configurations which produced the lowest value for AICc (i.e., AICc=0;
+  following Kass et al.~(2021)).
+\end{enumerate}
+
+\hypertarget{model-visualisation}{%
+\subsubsection{2.5. Model visualisation}\label{model-visualisation}}
+
+\begin{itemize}
+\tightlist
+\item
+  Need to discuss how we map the different rasters, and maybe how we
+  quantified the difference in suitability projections between climate
+  layers?
+\end{itemize}
+
+All modelling and statistical analyses were conducted in R ver. 4.0.3 (R
+Core Team, 2020). All values presented in text are presentedas mean ±
+standard error, unless otherwise stated. A standardised ODMAP methods
+protocol (Overview, Data, Model, Assessment, and Prediction) has been
+completed for this study and can be found in Supplementary File S1.
+ODMAP standardises the reporting of SDM modelling studies to improve
+transparency and reproducibility (Zurell et al., 2020).
+
+\hypertarget{results}{%
+\section{3. Results}\label{results}}
+
+\hypertarget{discussion}{%
+\section{4. Discussion}\label{discussion}}
+
+\hypertarget{declaration-of-competing-interest}{%
+\section{Declaration of Competing
+Interest}\label{declaration-of-competing-interest}}
+
+The authors declare that they have no known competing financial
+interests or personal relationships that could have appeared to
+influence the work reported in this paper.
+
+\hypertarget{data-availability}{%
+\section{Data availability}\label{data-availability}}
+
+All data and code required to reproduce the analyses are available in a
+public GitHub repository:
+https://github.com/guysutton/MS\_climate\_variable\_selection\_sdm
+
+\hypertarget{acknlowedgements}{%
+\section{Acknlowedgements}\label{acknlowedgements}}
+
+CVS and GFS acknowledge funding from the South African Working for Water
+(WfW) programme of the Department of Forestry, Fisheries and the
+Environment: Natural Resource Management Programmes (DFFE: NRMP).
+Funding was also provided by the South African Research Chairs
+Initiative of the Department of Science and Technology and the National
+Research Foundation (NRF) of South Africa. Any opinions, finding,
+conclusions or recommendations expressed in this material are those of
+the authors and the NRF does not accept any liability in this regard.
+
+\hypertarget{supplementary-materials}{%
+\section{Supplementary materials}\label{supplementary-materials}}
+
+\renewcommand\refname{References}
   \bibliography{bibliography.bib}

### 2023-06-05 manuscript/manuscript.qmd
Add spatial autocorrelation text in methods

@@
 keywords: 
-  - keyword1
-  - keyword2
+  - Species distribution model
+  - Variable selection
+  - Ecological model
+  - WORLDCLIM
 date: last-modified
@@
 
+- Clarke to add information on where we sourced GPS records - link to key papers, maybe a supplementary table (if required).
+- How many records do we have?
+
+Spatial autocorrelation is an important factor that may affect model outputs.
+Filtering of species occurrence data may limit the inherent biases in the data and improve model quality (Veloz, 2009).
+To avoid pseudo-replication, only one occurrencenrecord per 2.5 min grid cell was used for model calibration.
+Species occurrence datasets were thinned using the ‘*spThin*’ package [@Aiello-Lammens2015], and spatial autocorrelation analyses were performed using the ‘*ecospat*’ package [@DiCola2017].
+Eighty-eight occurrence records were retained to calibrate native-range models.
+Spatial thinning was also performed for the invaded range occurrences, however, there was no evidence for spatial autocorrelation.
+As such, no occurrence records were thinned for the invaded-range dataset.
+
 ### 2.2. Environmental predictors
@@
 Climate data were obtained by downloading the standard set of 19 bioclimatic variables from the WorldClim ver.
-2.1 database (Fick and Hijmans, 2017) (data available at: www.worldclim.org/download.
-html).
-This dataset is representative of annual and seasonal means and variation of temperature and precipitation metrics averaged over the 1950–2000 time period (current climate) at a 2.5 min resolution.
-These variables have been shown to effectively model the climatic suitability for non-native insects (e.g., Trethowan et al., 2011).
+2.1 database [@Fick2017] (data available at: www.worldclim.org/download.html).
+This dataset is representative of annual and seasonal means and variation of temperature and precipitation metrics averaged over the 1950–2000 time period (current climate) at a 2.5 minute resolution.
+These variables have been shown to effectively model the climatic suitability for non-native insects (e.g., @Sutton2022).
 
@@
 
-MaxEnt (ver. 3.4.3) was implemented in the ‘ dismo’ R package (Hijmans et al., 2021 ).
+MaxEnt (ver. 3.4.3) was implemented in the ‘*dismo*’ R package (@Hijmans2017).
 

### 2023-06-05 manuscript/manuscript.qmd
Bulk dump of text to intro and methods

@@
       model: 3p
-      layout: twocolumn
+      layout: onecolumn
       cite-style: authoryear
@@
 SDM’s typically take the form of correlative or mechanistic models that correlate species presence/absences (or pseudo-absences) to environmental covariates to identify suitable climatic conditions for the study taxon [@Elith2011].
-The Maximum Entropy species distribution model (hereafter ‘MaxEnt’) is amongst the most popular methods for climate modelling studies and has been shown to perform well compared to alternative modelling algorithms [@Wisz2008; @Phillips2017].
+The Maximum Entropy species distribution model (hereafter ‘*MaxEnt*’) is amongst the most popular methods for climate modelling studies and has been shown to perform well compared to alternative modelling algorithms [@Wisz2008; @Phillips2017].
 It uses maximum entropy to distinguish between environmental conditions where the focal taxon is present from environmental conditions at sites without confirmed presence records for the taxon [@Elith2011].
@@
 
-**Next paragraphs**
+The methods used for selecting covariates for SDM studies typically fall under four categories: (1) creating a set of uncorrelated covariates from a larger expert-chosen group of 15-30 covariates, using correlation analyses and/or variance inflation factors (VIF's) [@Dormann2013], (2) using principal component analysis (PCA) to reduce dimensionality of a set of covariates [@Junior2018], (3) using expert-opinion to select covariates that may be relevant based on the knowledge of the focal species biology, physiology and ecological requirements [@Petitpierre2017; @Scherrer2019], and more recently (4) automated statistical model-based covariate selection techniques, such as '*embedded methods*', e.g.
+RIDGE and LASSO [@Guisan2002; @Guyon2003; @Saeys2007; @Adde2023].
+For a more detailed overview of the different variable selection methods, we direct readers to several recent reviews on this topic [@Fois2018; @Fourcade2018; @Melo-Merino2020].
 
-- Review current methods used to select variables, e.g. R2, VIF, PCA, expert opinion, and most recently, automated selection / model-based selection.
+While there are guidelines for how to implement these methods individually, e.g.
+using an |*r*| > 0.70 or a VIF > 2-5 as a threshold for removing significantly correlated predictors [@Dormann2013], there is currently no consensus on which method to use for a particular study (but see @Adde2023).
+As such, we aimed to demonstrate the variation in climatic suitability raster projections and predictive accuracy of MaxEnt models calibrated using the different covariate selection methods, and discuss the potential implications for model inference and possible management programmes based on these inferences.
+To demonstrate, we modelled the potential climatic suitability (i.e.
+the climatic similarity relative to the climate space occupied in the model calibration region) for the invasive insect pest *Diaphorina citri* Kuwayama (Asian citrus psyllid) (Hemiptera: Psyllidae), that is native to Asia but has become invasive in many non-native regions around the world, such as the USA, Brazil and Africa ( **Insert citations**, **Insert a map of global distribution?** ).
 
-- Aims of this paper are to show how variable selection can influence model outputs and inferences.
+- Clarke: Can you please insert 1-2 sentences about the impact of ACP and its association with HLB?
 
@@
 
-### 2.1. Species occurrence records
+### Species occurrence records
 
@@
 Spatial autocorrelation is an important factor that may affect model outputs.
-Filtering of species occurrence data may limit the inherent biases in the data and improve model quality (Veloz, 2009).
-To avoid pseudo-replication, only one occurrencenrecord per 2.5 min grid cell was used for model calibration.
+Filtering of species occurrence data may limit the inherent biases in the data and improve model quality [@Veloz2009].
+To avoid pseudo-replication, only one occurrence record per 2.5 minute grid cell was used for model calibration.
 Species occurrence datasets were thinned using the ‘*spThin*’ package [@Aiello-Lammens2015], and spatial autocorrelation analyses were performed using the ‘*ecospat*’ package [@DiCola2017].
@@
 
-### 2.2. Environmental predictors
+### Environmental predictors
 
@@
 
-### 2.3. Model calibration
+### Model calibration
 
@@
 Given that MaxEnt is a presence/pseudo-absence modelling algorithm, model calibration requires a user-defined geographic background to sample the climate of representative grid cells where the focal species is assumed to be absent (i.e., background points or pseudo-absences).
-Background definition can have a significant effect on model output (VanDerWal et al., 2009).
-The background should ideally represent the geographic areas available to the focal species, omitting areas where species absence is due to historical factors, dispersal constraints and/or biotic interactions (Sanin and Anderson, 2018).
-Following Webber et al.
-(2011), we defined the model background using the Koppen-Geiger climate classification system (available at: http://koeppen-geiger.vu-w ien.ac.at).
+Background definition can have a significant effect on model output [@VanDerWal2009].
+The background should ideally represent the geographic areas available to the focal species, omitting areas where species absence is due to historical factors, dispersal constraints and/or biotic interactions [@Sanin2018].
+We defined the model background using the Koppen-Geiger climate classification system (available at: http://koeppen-geiger.vu-wien.ac.at) [@Webber2011].
 Only Koppen-Geiger climate zones that contained at least one native-range occurrence record for D.
@@
 1a).
+
+* Need to update the below with appropriate references to 'terra' package
+
 The Koppen-Geiger climate zones were intersected with the occurrence records using the ‘ raster’ R package (Hijmans, 2022).
@@
 
-MaxEnt models were parameterised with default settings for multiple parameters, including: convergence = 105, maximum number of iterations = 500 and prevalence = 0.5.
-The ‘fade by clamping’ option was selected to prevent extrapolation well outside the range of climatic values in the model training area (Philips et al., 2017 ).
+MaxEnt models were specified with default settings for multiple parameters, including: convergence = 105, maximum number of iterations = 500 and prevalence = 0.5.
+The ‘fade by clamping’ option was selected to prevent extrapolation well outside the range of climatic values in the model training area [@Phillips2017].
 Model predictions were obtained using the ‘logistic output’ to create continuous climatic suitability raster layers scaled between 0 (climatically unsuitable) and 1 (climatically suitable).
 
-### 2.4. Model evaluation
+### Model evaluation
 
@@
 
-### 2.5. Model visualisation
+### Model visualisation
 
@@
 
-All modelling and statistical analyses were conducted in R ver.
-4.0.3 (R Core Team, 2020).
-All values presented in text are presentedas mean ± standard error, unless otherwise stated.
+All modelling and statistical analyses were conducted in *R* ver.
+4.3.0 [@Team2023].
+All values presented in text are presented as mean ± standard error, unless otherwise stated.
 A standardised ODMAP methods protocol (Overview, Data, Model, Assessment, and Prediction) has been completed for this study and can be found in Supplementary File S1.

### 2023-06-05 manuscript/manuscript.tex
Bulk dump of text to intro and methods

@@
   3p,
-  twocolumn]{elsarticle}
+  onecolumn]{elsarticle}
 
@@
 \makeatother
-\usepackage{float}
-\makeatletter
-\let\oldlt\longtable
-\let\endoldlt\endlongtable
-\def\longtable{\@ifnextchar[\longtable@i \longtable@ii}
-\def\longtable@i[#1]{\begin{figure}[H]
-\onecolumn
-\begin{minipage}{0.5\textwidth}
-\oldlt[#1]
-}
-\def\longtable@ii{\begin{figure}[H]
-\onecolumn
-\begin{minipage}{0.5\textwidth}
-\oldlt
-}
-\def\endlongtable{\endoldlt
-\end{minipage}
-\twocolumn
-\end{figure}}
-\makeatother
 \journal{Ecological Modelling}
@@
   pdfauthor={Clarke van Steenderen; Guy F. Sutton},
-  pdfkeywords={keyword1, keyword2},
+  pdfkeywords={Species distribution model, Variable
+selection, Ecological model, WORLDCLIM},
   colorlinks=true,
@@
 \begin{keyword}
-    keyword1 \sep 
-    keyword2
+    Species distribution model \sep Variable selection \sep Ecological
+model \sep 
+    WORLDCLIM
 \end{keyword}
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, interior hidden, sharp corners, frame hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, boxrule=0pt, enhanced, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 the study taxon \citep{Elith2011}. The Maximum Entropy species
-distribution model (hereafter `MaxEnt') is amongst the most popular
-methods for climate modelling studies and has been shown to perform well
-compared to alternative modelling algorithms
+distribution model (hereafter `\emph{MaxEnt}') is amongst the most
+popular methods for climate modelling studies and has been shown to
+perform well compared to alternative modelling algorithms
 \citep{Wisz2008, Phillips2017}. It uses maximum entropy to distinguish
@@
 
-\textbf{Next paragraphs}
+The methods used for selecting covariates for SDM studies typically fall
+under four categories: (1) creating a set of uncorrelated covariates
+from a larger expert-chosen group of 15-30 covariates, using correlation
+analyses and/or variance inflation factors (VIF's) \citep{Dormann2013},
+(2) using principal component analysis (PCA) to reduce dimensionality of
+a set of covariates \citep{Junior2018}, (3) using expert-opinion to
+select covariates that may be relevant based on the knowledge of the
+focal species biology, physiology and ecological requirements
+\citep{Petitpierre2017, Scherrer2019}, and more recently (4) automated
+statistical model-based covariate selection techniques, such as
+`\emph{embedded methods}', e.g.~RIDGE and LASSO
+\citep{Guisan2002, Guyon2003, Saeys2007, Adde2023}. For a more detailed
+overview of the different variable selection methods, we direct readers
+to several recent reviews on this topic
+\citep{Fois2018, Fourcade2018, Melo-Merino2020}.
+
+While there are guidelines for how to implement these methods
+individually, e.g.~using an \textbar{}\emph{r}\textbar{} \textgreater{}
+0.70 or a VIF \textgreater{} 2-5 as a threshold for removing
+significantly correlated predictors \citep{Dormann2013}, there is
+currently no consensus on which method to use for a particular study
+(but see \citet{Adde2023}). As such, we aimed to demonstrate the
+variation in climatic suitability raster projections and predictive
+accuracy of MaxEnt models calibrated using the different covariate
+selection methods, and discuss the potential implications for model
+inference and possible management programmes based on these inferences.
+To demonstrate, we modelled the potential climatic suitability (i.e.~the
+climatic similarity relative to the climate space occupied in the model
+calibration region) for the invasive insect pest \emph{Diaphorina citri}
+Kuwayama (Asian citrus psyllid) (Hemiptera: Psyllidae), that is native
+to Asia but has become invasive in many non-native regions around the
+world, such as the USA, Brazil and Africa ( \textbf{Insert citations},
+\textbf{Insert a map of global distribution?} ).
 
 \begin{itemize}
+\tightlist
 \item
-  Review current methods used to select variables, e.g.~R2, VIF, PCA,
-  expert opinion, and most recently, automated selection / model-based
-  selection.
-\item
-  Aims of this paper are to show how variable selection can influence
-  model outputs and inferences.
+  Clarke: Can you please insert 1-2 sentences about the impact of ACP
+  and its association with HLB?
 \end{itemize}
@@
 \hypertarget{species-occurrence-records}{%
-\subsubsection{2.1. Species occurrence
+\subsubsection{Species occurrence
 records}\label{species-occurrence-records}}
 
+\begin{itemize}
+\tightlist
+\item
+  Clarke to add information on where we sourced GPS records - link to
+  key papers, maybe a supplementary table (if required).
+\item
+  How many records do we have?
+\end{itemize}
+
+Spatial autocorrelation is an important factor that may affect model
+outputs. Filtering of species occurrence data may limit the inherent
+biases in the data and improve model quality \citep{Veloz2009}. To avoid
+pseudo-replication, only one occurrence record per 2.5 minute grid cell
+was used for model calibration. Species occurrence datasets were thinned
+using the `\emph{spThin}' package \citep{Aiello-Lammens2015}, and
+spatial autocorrelation analyses were performed using the
+`\emph{ecospat}' package \citep{DiCola2017}. Eighty-eight occurrence
+records were retained to calibrate native-range models. Spatial thinning
+was also performed for the invaded range occurrences, however, there was
+no evidence for spatial autocorrelation. As such, no occurrence records
+were thinned for the invaded-range dataset.
+
 \hypertarget{environmental-predictors}{%
-\subsubsection{2.2. Environmental
+\subsubsection{Environmental
 predictors}\label{environmental-predictors}}
@@
 Climate data were obtained by downloading the standard set of 19
-bioclimatic variables from the WorldClim ver. 2.1 database (Fick and
-Hijmans, 2017) (data available at: www.worldclim.org/download. html).
+bioclimatic variables from the WorldClim ver. 2.1 database
+\citep{Fick2017} (data available at: www.worldclim.org/download.html).
 This dataset is representative of annual and seasonal means and
 variation of temperature and precipitation metrics averaged over the
-1950--2000 time period (current climate) at a 2.5 min resolution. These
-variables have been shown to effectively model the climatic suitability
-for non-native insects (e.g., Trethowan et al., 2011).
+1950--2000 time period (current climate) at a 2.5 minute resolution.
+These variables have been shown to effectively model the climatic
+suitability for non-native insects (e.g., \citet{Sutton2022}).
 
 \hypertarget{model-calibration}{%
-\subsubsection{2.3. Model calibration}\label{model-calibration}}
+\subsubsection{Model calibration}\label{model-calibration}}
 
-MaxEnt (ver. 3.4.3) was implemented in the ` dismo' R package (Hijmans
-et al., 2021 ).
+MaxEnt (ver. 3.4.3) was implemented in the `\emph{dismo}' R package
+(\citet{Hijmans2017}).
 
@@
 Background definition can have a significant effect on model output
-(VanDerWal et al., 2009). The background should ideally represent the
+\citep{VanDerWal2009}. The background should ideally represent the
 geographic areas available to the focal species, omitting areas where
 species absence is due to historical factors, dispersal constraints
-and/or biotic interactions (Sanin and Anderson, 2018). Following Webber
-et al.~(2011), we defined the model background using the Koppen-Geiger
-climate classification system (available at: http://koeppen-geiger.vu-w
-ien.ac.at). Only Koppen-Geiger climate zones that contained at least one
+and/or biotic interactions \citep{Sanin2018}. We defined the model
+background using the Koppen-Geiger climate classification system
+(available at: http://koeppen-geiger.vu-wien.ac.at) \citep{Webber2011}.
+Only Koppen-Geiger climate zones that contained at least one
 native-range occurrence record for D. rubiformis in Australia were used
 as the background area from which background points were drawn for model
-calibration (Fig. 1a). The Koppen-Geiger climate zones were intersected
-with the occurrence records using the ` raster' R package (Hijmans,
-2022). We randomly sampled 10 000 points (the default number used for
-Maxent; Merow et al.~(2013) ) from within this background definition
-using the `dismo' R package (Hijmans et al., 2021).
+calibration (Fig. 1a).
 
-MaxEnt models were parameterised with default settings for multiple
+\begin{itemize}
+\tightlist
+\item
+  Need to update the below with appropriate references to `terra'
+  package
+\end{itemize}
+
+The Koppen-Geiger climate zones were intersected with the occurrence
+records using the ` raster' R package (Hijmans, 2022). We randomly
+sampled 10 000 points (the default number used for Maxent; Merow et
+al.~(2013) ) from within this background definition using the `dismo' R
+package (Hijmans et al., 2021).
+
+MaxEnt models were specified with default settings for multiple
 parameters, including: convergence = 105, maximum number of iterations =
@@
 prevent extrapolation well outside the range of climatic values in the
-model training area (Philips et al., 2017 ). Model predictions were
+model training area \citep{Phillips2017}. Model predictions were
 obtained using the `logistic output' to create continuous climatic
@@
 \hypertarget{model-evaluation}{%
-\subsubsection{2.4. Model evaluation}\label{model-evaluation}}
+\subsubsection{Model evaluation}\label{model-evaluation}}
 
@@
 \hypertarget{model-visualisation}{%
-\subsubsection{2.5. Model visualisation}\label{model-visualisation}}
+\subsubsection{Model visualisation}\label{model-visualisation}}
 
@@
 
-All modelling and statistical analyses were conducted in R ver. 4.0.3 (R
-Core Team, 2020). All values presented in text are presentedas mean ±
-standard error, unless otherwise stated. A standardised ODMAP methods
-protocol (Overview, Data, Model, Assessment, and Prediction) has been
-completed for this study and can be found in Supplementary File S1.
+All modelling and statistical analyses were conducted in \emph{R} ver.
+4.3.0 \citep{Team2023}. All values presented in text are presented as
+mean ± standard error, unless otherwise stated. A standardised ODMAP
+methods protocol (Overview, Data, Model, Assessment, and Prediction) has
+been completed for this study and can be found in Supplementary File S1.
 ODMAP standardises the reporting of SDM modelling studies to improve

### 2023-06-05 manuscript/manuscript.tex
Correct section header formatting

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, boxrule=0pt, enhanced, interior hidden]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, interior hidden, enhanced, boxrule=0pt, breakable, frame hidden]}{\end{tcolorbox}}\fi
 
 \hypertarget{introduction}{%
-\section{1. Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}}
 
@@
 \hypertarget{methods-and-materials}{%
-\section{2. Methods and materials}\label{methods-and-materials}}
+\section{Methods and materials}\label{methods-and-materials}}
 
 \hypertarget{species-occurrence-records}{%
-\subsubsection{Species occurrence
+\subsection{Species occurrence
 records}\label{species-occurrence-records}}
@@
 \hypertarget{environmental-predictors}{%
-\subsubsection{Environmental
-predictors}\label{environmental-predictors}}
+\subsection{Environmental predictors}\label{environmental-predictors}}
 
@@
 \hypertarget{model-calibration}{%
-\subsubsection{Model calibration}\label{model-calibration}}
+\subsection{Model calibration}\label{model-calibration}}
 
@@
 \hypertarget{model-evaluation}{%
-\subsubsection{Model evaluation}\label{model-evaluation}}
+\subsection{Model evaluation}\label{model-evaluation}}
 
@@
 \hypertarget{model-visualisation}{%
-\subsubsection{Model visualisation}\label{model-visualisation}}
+\subsection{Model visualisation}\label{model-visualisation}}
 
@@
 \hypertarget{results}{%
-\section{3. Results}\label{results}}
+\section{Results}\label{results}}
 
 \hypertarget{discussion}{%
-\section{4. Discussion}\label{discussion}}
+\section{Discussion}\label{discussion}}
 
